---
jupyter: ir
---

```{r}
#| vscode: {languageId: r}
df = read.csv('data_final.csv')
print(head(df))
```

```{r}
#| vscode: {languageId: r}
summary(df)
```

```{r}
#| vscode: {languageId: r}
unique_region <- unique(df$WHO.Region)
print(unique_region)
```

```{r}
#| vscode: {languageId: r}
unique_iso <- unique(df$ISO3)
print(unique_iso)
```

```{r}
#| vscode: {languageId: r}
unique_country <- unique(df$WHO.Country.Name)
print(unique_country)
```

Average.NO2 is Average NO2 for country

2010, 2014 to 2019


```{r}
#| vscode: {languageId: r}
library(dplyr)

# Balancing data choosing countries with data for years 2010, 2016, 2017, and 2018
years_to_keep <- c(2010, 2016, 2017, 2018)

# balanced data
years_to_keep <- c(2010, 2014, 2015, 2016, 2017, 2018)

# Filter the data frame
filtered_df <- df %>%
  group_by(WHO.Country.Name) %>%
  filter(all(years_to_keep %in% Measurement.Year))

# Display the filtered data frame
print(head(filtered_df))
```

```{r}
#| vscode: {languageId: r}
library(ggplot2)
print(names(df))

#hist(df$NO2....g.m3.)

df %>%
  ggplot( aes(x = WHO.Region, y = NO2..μg.m3., fill=WHO.Region)) +
  geom_boxplot() +ggtitle("Comparison of NO2 between WHO Region")

hist(df$Population)

hist(df$NO2.temporal.coverage....)

```

```{r}
#| vscode: {languageId: r}

df1 <- subset(df, select=c(Measurement.Year, WHO.Country.Name))

df2 <- df1[order(df1$Measurement.Year), ]
df2

table <- table(df1)

# Create a new data frame with counts for each combination of Year and Country
count_df <- df1 %>%
  group_by(Measurement.Year, WHO.Country.Name) %>%
  summarize(Frequency = n())

# Display the resulting data frame
print(count_df)

count_df[count_df$WHO.Country.Name=="Korea",]
```


```{r}

df_region <- subset(df, select=c(Measurement.Year, WHO.Region))


# Create a new data frame with counts for each combination of Year and Country
df_region <- df_region %>%
  group_by(Measurement.Year, WHO.Region) %>%
  summarize(Frequency = n())

# Display the resulting data frame
print(df_region)
```



EUR vs. SEA

```{r}


sea <- subset(df, select=NO2..μg.m3.,
                      subset=WHO.Region=="South East Asia Region", drop=T)

eur  <- subset(df, select=NO2..μg.m3.,
                      subset=WHO.Region=="European Region", drop=T)

t.test(sea, eur, alt="greater")

```

1. INCUDE IN FINAL PROJECT AND PRESENTATION

## Does South East Asia Region have more NO2 emissions than the European Region in 2019?

Fail to reject the null hypothesis: The NO2 emissions in the South East Asia Region and European Region are similar.   
Reject the null hypothesis or alternative hypothesis: The NO2 emissions in the South East Asia Region is greater than in the European Region.  


```{r}

df2019 <- df[df$Measurement.Year==2019, ]

sea2 <- subset(df2019, select=NO2..μg.m3.,
                      subset=WHO.Region=="South East Asia Region", drop=T)

eur2  <- subset(df2019, select=NO2..μg.m3.,
                      subset=WHO.Region=="European Region", drop=T)

t.test(sea2, eur2, alt="greater")

```

EUR vs. SEA in 2019
repeating test with Average NO2


```{r}
sea2 <- subset(df2019, select=Average.NO2,
                      subset=WHO.Region=="South East Asia Region", drop=T)

eur2  <- subset(df2019, select=Average.NO2,
                      subset=WHO.Region=="European Region", drop=T)

t.test(sea2, eur2, alt="greater")
```

1. INCLUDE IN FINAL PROJECT 

Additionally, a bootstrap test is performed for ratio of means of “NO2.” The 95% bootstrap percentile interval for the ratio of means is (0.6566061 1.7743924). The estimated bias is 0.03308387. The bootstrap tests also shows there is 95% confidence in the alternative hypothesis.

Note: adjust and control for years

```{r}
sea3 <- df2019[df2019$WHO.Region == "South East Asia Region", ]
eur3 <- df2019[df2019$WHO.Region == "European Region", ]
N <- 100000

ratiomeans<- numeric(N)

for (i in 1:N)
{
  sea_sample <- sample(df$NO2..μg.m3.[df$WHO.Region == 'South East Asia Region'], length(sea3), replace = TRUE)
  eur_sample <- sample(df$NO2..μg.m3.[df$WHO.Region == 'European Region'], length(eur3), replace = TRUE)
  ratiomeans[i] <- mean(sea_sample)/mean(eur_sample)
}

quantile(ratiomeans, c(0.025, 0.975))
```

Calculate bias
```{r}
sea_mean <- mean(df$NO2..μg.m3.[df$WHO.Region == 'South East Asia Region'])
eur_mean <- mean(df$NO2..μg.m3.[df$WHO.Region == 'European Region'])
ratio_actual <- sea_mean/eur_mean
(bias= mean(ratiomeans)-mean(ratio_actual)) #estimated bias
```


According to the existing WHO air quality guidelines, an annual average indoor nitrogen dioxide guideline of 40 μg/m3 is recommended. In the WHO dataset, a new column was added labeling NO2 above and within 40 μg/m3. Chi-squared test is performed to assess 

Fail to reject the null hypothesis: The compliance with recommended NO2 of regions are independent. The compliance with recommended NO2 between the regions are not different.

Reject the null hypothesis or alternative hypothesis: The compliance with recommended NO2 are dependent on the region. The compliance with recommended NO2 are different between regions.

"An annual average indoor nitrogen dioxide guideline of 40 μg/m3, consistent with the existing WHO air quality guideline, is recommended."
https://www.ncbi.nlm.nih.gov/books/NBK138707/


```{r}
df$WHO.Guideline <- ifelse(df$NO2..μg.m3. >= 41.0, "above", "within")
print(head(df))

df_region_chi <- df[!is.na(df$WHO.Region) & !is.na(df$WHO.Guideline),]

tableRegion <- table(df$WHO.Region, df$WHO.Guideline)
tableRegion

chisq.test(tableRegion)

```



```{r}

germany1 <- subset(df, select=NO2..μg.m3.,
                      subset=WHO.Country.Name=="Germany", drop=T)

france1  <- subset(df, select=NO2..μg.m3.,
                      subset=WHO.Country.Name=="France", drop=T)

t.test(germany1, france1, alt="greater")

```


```{r}
df2010 <- df[df$Measurement.Year==2010, ]

germany2 <- subset(df2010, select=NO2..μg.m3.,
                      subset=WHO.Country.Name=="Germany", drop=T)

france2  <- subset(df2010, select=NO2..μg.m3.,
                      subset=WHO.Country.Name=="France", drop=T)

t.test(france2, germany2, alt="less")
```



```{r}

germany3 <- subset(df, select=Population,
                      subset=WHO.Country.Name=="Germany", drop=T)

france3  <- subset(df, select=Population,
                      subset=WHO.Country.Name=="France", drop=T)

t.test(germany3, france3, alt="greater")

```

```{r}

df_germany <- df[df$WHO.Country.Name=="Germany", ]

germany2016 <- subset(df_germany, select=NO2..μg.m3.,
                      subset=Measurement.Year=="2016", drop=T)

germany2018  <- subset(df_germany, select=NO2..μg.m3.,
                      subset=Measurement.Year=="2018", drop=T)

t.test(germany2018, germany2016, alt="less")


```




```{r}

df_france <- df[df$WHO.Country.Name=="France", ]

france2010 <- subset(df_france, select=NO2..μg.m3.,
                      subset=Measurement.Year=="2010", drop=T)

france2016  <- subset(df_france, select=NO2..μg.m3.,
                      subset=Measurement.Year=="2016", drop=T)

t.test(france2016, france2010, alt="less")


```
2a. INCLUDE IN FINAL PRESENTATION AND PROJECT

## Given that South East Asian has greater NO2 emissions than Europe, additional testing was performed on France and India. 

## Does France have less NO2 emissions in 2016 than in 2010? 

A bootstrap test is performed for ratio of means of “NO2.” The 95% bootstrap percentile interval for the ratio of means is (0.6566061 1.7743924). The estimated bias is approximately 0.03. The bootstrap tests also shows there is 95% confidence in the alternative hypothesis.

Yes, France has decreased their NO2 emissions in 2016.


```{r}
df_france_2010 <- df_france[df_france$Measurement.Year == "2010", ]
df_france_2016 <- df_france[df_france$Measurement.Year == "2016", ]
N <- 100000

ratiomeans<- numeric(N)

for (i in 1:N)
{
  fr_sample_2010 <- sample(df_france_2010$NO2..μg.m3.[df_france_2010$Measurement.Year == '2010'], length(df_france_2010), replace = TRUE)
  fr_sample_2016 <- sample(df_france_2016$NO2..μg.m3.[df_france_2016$Measurement.Year == '2016'], length(df_france_2016), replace = TRUE)
  ratiomeans[i] <- mean(fr_sample_2010)/mean(fr_sample_2016)
}

quantile(ratiomeans, c(0.025, 0.975))

fr_mean_2010 <- mean(df_france_2010$NO2..μg.m3.[df_france_2010$Measurement.Year == '2010'])
fr_mean_2016 <- mean(df_france_2016$NO2..μg.m3.[df_france_2016$Measurement.Year == '2016'])
ratio_actual <- fr_mean_2010/fr_mean_2016
bias= mean(ratiomeans)-mean(ratio_actual) #estimated bias

print(paste("The estimated bias is", bias))
```




```{r}

germany2013 <- subset(df_germany, select=NO2..μg.m3.,
                      subset=Measurement.Year=="2013", drop=T)

germany2019  <- subset(df_germany, select=NO2..μg.m3.,
                      subset=Measurement.Year=="2019", drop=T)

t.test(germany2019, germany2013, alt="less")
```

2b. INCLUDE IN FINAL PROJECT AND PRESENTATION
## Does India have less NO2 emissions in 2018 than in 2014? After performing a t-test, the p-value was slightly greater than 0.05, which fails to reject the null hypothesis.

India did not decrease NO2 emissions in 2018.

```{r}


df_india <- df[df$WHO.Country.Name=="India", ]

india2014 <- subset(df_germany, select=NO2..μg.m3.,
                      subset=Measurement.Year=="2014", drop=T)

india2018  <- subset(df_germany, select=NO2..μg.m3.,
                      subset=Measurement.Year=="2018", drop=T)

t.test(india2018, india2014, alt="less")


```

3. INCLUDE IN PRESENTATION

## Is global NO2 dependent on Year?

According to the existing WHO air quality guidelines, an annual average indoor nitrogen dioxide guideline of 40 μg/m3 is recommended. In the WHO dataset, a new column was added labeling NO2 above and within 40 μg/m3. Chi-squared test is performed to assess if compliance with recommended NO2 is dependent on year in the subset.

Fail to reject the null hypothesis: The compliance with recommended NO2 are independent from years. The compliance with recommended NO2 does not differ between years.

Reject the null hypothesis or alternative hypothesis: The compliance with recommended NO2 are dependent on the year. The compliance with recommended NO2 differs between the years.

The Chi-squared test result in p-value < 2.2e-16, confirming the alternative hypothesis. Therefore, compliance with recommended NO2 differs between the years. 

"An annual average indoor nitrogen dioxide guideline of 40 μg/m3, consistent with the existing WHO air quality guideline, is recommended."
https://www.ncbi.nlm.nih.gov/books/NBK138707/


```{r}

df_global <- df[(df$Measurement.Year>=2013 & df$Measurement.Year<=2019),]

table_year<- table(df_global$Measurement.Year, df_global$WHO.Guideline)
table_year

chisq.test(table_year)

```

To take a closer look a subset was limited to India, Germany, and France in between the date range of 2013 to 2019.

According to the existing WHO air quality guidelines, an annual average indoor nitrogen dioxide guideline of 40 μg/m3 is recommended. In the WHO dataset, a new column was added labeling NO2 above and within 40 μg/m3. Chi-squared test is performed to assess if compliance with recommended NO2 is dependent on year in the subset.

Fail to reject the null hypothesis: The compliance with recommended NO2 are independent from years. The compliance with recommended NO2 does not differ between years.

Reject the null hypothesis or alternative hypothesis: The compliance with recommended NO2 are dependent on the year. The compliance with recommended NO2 differs between the years.

"An annual average indoor nitrogen dioxide guideline of 40 μg/m3, consistent with the existing WHO air quality guideline, is recommended."
https://www.ncbi.nlm.nih.gov/books/NBK138707/


```{r}

df_global2 <- df_global[df_global$WHO.Country.Name=="India" | df_global$WHO.Country.Name=="Germany" | df_global$WHO.Country.Name=="France", ]

df_global2 <- df_global2[(df_global2$Measurement.Year>=2013 & df_global2$Measurement.Year<=2019),]

table_year2 <- table(df_global2$Measurement.Year, df_global2$WHO.Guideline)
table_year2

chisq.test(table_year2)

```

