---
title: "women_employment"
output: html_document
date: "2023-11-26"
---

read in output_data_v2.csv

```{r}
women <- read.csv('output_data_v2.csv')

print(colnames(women))

```

STEM vs. non-STEM  
- compare wage  
- compare employment  

Computer vs. non-Computer  
- compare wage  
- compare employment  

Industry SOC major groups classification  
- https://www.bls.gov/soc/2018/major_groups.htm  

Employment STEM classification  
- https://www.bls.gov/soc/Attachment_C_STEM_2018.pdf  

Compare top three industries/SOC majors  

https://www.bls.gov/spotlight/2017/science-technology-engineering-and-mathematics-stem-occupations-past-present-and-future/home.htm

```{r}

stem_t <- na.omit(women)
women_clean <- na.omit(women)

notstem <- subset(stem_t, select=Percent.women.2022,
                      subset=STEM=="no", drop=T)

stem  <- subset(stem_t, select=Percent.women.2022,
                      subset=STEM=="yes", drop=T)

t.test(notstem, stem,alt="greater")

dist_stem <- women_clean$Percent.women.2022[women_clean$STEM == "yes"]
print(dist_stem)
hist(dist_stem)

dist_nostem <- women_clean$Percent.women.2022[women_clean$STEM == "no"]
print(dist_nostem)
hist(dist_nostem)
```

Bootsrap percent women STEM vs. Non-Stem

```{r}



stem_boot <- women_clean[women_clean$STEM == "yes", ]
nostem_boot <- women_clean[women_clean$STEM == "no", ]
N <- 100000

ratiomeans<- numeric(N)

for (i in 1:N)
{
  stem_sample <- sample(women_clean$Percent.women.2022[women_clean$STEM == 'yes'], length(stem_boot), replace = TRUE)
  nostem_sample <- sample(women_clean$Percent.women.2022[women_clean$STEM == 'no'], length(nostem_boot), replace = TRUE)
  ratiomeans[i] <- mean(stem_sample)/mean(nostem_sample)
}

quantile(ratiomeans, c(0.025, 0.975))

hist(ratiomeans)

stem_mean <- mean(women_clean$Percent.women.2022[women_clean$STEM == 'yes'])
nostem_mean <- mean(women_clean$Percent.women.2022[women_clean$STEM == 'no'])
ratio_actual <- stem_mean/nostem_mean
(bias= mean(ratiomeans)-mean(ratio_actual)) #estimated bias


```



```{r}
stem_wage <- subset(women, select=Median.annual.wage.2022,
                      subset=STEM=="yes", drop=T)

nostem_wage <- subset(women, select=Median.annual.wage.2022,
                      subset=STEM=="no", drop=T)

t.test(stem_wage, nostem_wage, alt="greater")

```



```{r}

notcomputer <- subset(women, select=Percent.women.2022,
                      subset=Industry_SOCmajor!="Computer and Mathematical Occupations", drop=T)

computer <- subset(women, select=Percent.women.2022,
                      subset=Industry_SOCmajor=="Computer and Mathematical Occupations", drop=T)


t.test(notcomputer, computer, alt="greater")

```

```{r}

computer_wage <- subset(women, select=Median.annual.wage.2022,
                      subset=Industry_SOCmajor=="Computer and Mathematical Occupations", drop=T)

notcomputer_wage <- subset(women, select=Median.annual.wage.2022,
                      subset=Industry_SOCmajor!="Computer and Mathematical Occupations", drop=T)

t.test(computer_wage, notcomputer_wage, alt="greater")
```

Management Occupations
Legal Occupations
Computer and Mathematical Occupations
Healthcare Practitioners and Technical Occupations

employment men, employment women

```{r}
library(kableExtra)
library(dplyr)
library(tidyr)
topemployment <- women[(women$Industry_SOCmajor=="Management Occupations" |
                           women$Industry_SOCmajor=="Legal Occupations" |
                           women$Industry_SOCmajor=="Computer and Mathematical Occupations" |
                           women$Industry_SOCmajor=="Healthcare Practitioners and Technical Occupations"),]

#remove rows with NA
topemployment <- na.omit(topemployment)
topemployment_women <- topemployment[, c('Industry_SOCmajor', 'Employment..2022', 'Men.employment..2022')]


topemployment <- topemployment %>%
  group_by(Industry_SOCmajor) %>%
  summarise(women_employment = sum(Employment..2022), men_employment = sum(Men.employment..2022))

topemployment[, sapply(topemployment, is.numeric)] <- round(topemployment[, sapply(topemployment, is.numeric)], 0)

topemployment

topemployment <- topemployment[, c("women_employment", "men_employment")]

chisq.test(topemployment)


```

Top 4 Earning Employment Stem vs. Non-stem


```{r}

library(dplyr)
library(tidyr)


stem_chi <- women[, c('STEM', 'Employment..2022', 'Men.employment..2022')]

#remove rows with NA
stem_chi <- na.omit(stem_chi)

stem_chi <- stem_chi %>%
  group_by(STEM) %>%
  summarise(women_employment = sum(Employment..2022), men_employment = sum(Men.employment..2022))

stem_chi[, sapply(stem_chi, is.numeric)] <- round(stem_chi[, sapply(stem_chi, is.numeric)], 0)

stem_chi

stem_chi <- stem_chi[, c("women_employment", "men_employment")]

chisq.test(stem_chi)


```

